---
title: "Martin Tuohy - quarto exercise 5"
format: 
  html:
    output-file: "index.html"
---

```{r}
#| echo: true
#| warning: false
library(dplyr)
library(tidyr)
library(reactable)
```

1. Read in the CSV of Times Higher Education World University Ranking.

```{r}
df <- read.csv("http://manctsui.github.io/datasets/timesData.csv")
reactable(df, defaultPageSize = 4, searchable = TRUE)
```

2. Determine the rankings through the years of the undergraduate alma mater of the instructor for this course. 

```{r}
#| code-fold: true
#| warning: false
problem2 <- df |> 
  filter(grepl("University of California, Los Angeles", university_name))

reactable(problem2)
```

3. Plot the instructor’s undergraduate alma mater as year vs university rank

4. Repeat Problem 3 for UC Berkeley

::: {#fig-university-rank}
```{r}
#| code-fold: true
#| warning: false
x = problem2$year
y = problem2$world_rank

plot(x,y,
    main = "UCLA World Rank vs. Year",
    xlab = "Years", 
    ylab = "World Rank", 
    type="l" 
     )

problem4 <- df |> 
  filter(grepl("University of California, Berkeley", university_name))

x = problem4$year
y = problem4$world_rank

plot(x,y,
    main = "UC Berkeley World Rank vs. Year",
    xlab = "Years", 
    ylab = "World Rank", 
    type="l" 
     )
```
:::

5. Did the number of US universities that rank in the top 50 change much through the years?

:::{#fig-top-50-universities}
```{r}
#| code-fold: true
#| warning: false

df$world_rank <- as.numeric(as.character(df$world_rank))

problem5_1 <- df |>
  filter(world_rank <= 50 & year == 2011) |>
  arrange(world_rank) |>
  select("world_rank","university_name","year") 

reactable(problem5_1)

problem5_2 <- df |>
  filter(world_rank <= 50 & year == 2012) |>
  arrange(world_rank) |>
  select("world_rank","university_name","year") 

reactable(problem5_2)

problem5_3 <- df |>
  filter(world_rank <= 50 & year == 2013) |>
  arrange(world_rank) |>
  select("world_rank","university_name","year") 

reactable(problem5_3)

problem5_4 <- df |>
  filter(world_rank <= 50 & year == 2014) |>
  arrange(world_rank) |>
  select("world_rank","university_name","year") 

reactable(problem5_4)

problem5_5 <- df |>
  filter(world_rank <= 50 & year == 2015) |>
  arrange(world_rank) |>
  select("world_rank","university_name","year") 

reactable(problem5_5)

problem5_6 <- df |>
  filter(world_rank <= 50 & year == 2016) |>
  arrange(world_rank) |>
  select("world_rank","university_name","year") 

reactable(problem5_6)
```
:::